---
title: "Week 2"
subtitle: "Linear programming primal, dual, and sensitivity analysis"
author: "Gareth Green"
output: slidy_presentation
---


```{r echo = FALSE}
# Course: 5260 Math models for decision making
# Title: Week 2-Linear programming primal, dual, and sensitivity analysis
# Purpose: Basic demonstration of primal and dual with lpSolveAPI
# Date: March 17, 2018
# Author: Gareth Green

```

```{r echo = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

```


Set up boots and shoes problem
=============================================

<div style="float: left; width: 50%;">

```{r message = FALSE, warning = FALSE}
# Load packages
library(lpSolveAPI)

```

```{r}
# Define parameters of the lp
  ## Note, this set up shows 0 constraints and 2 variables
  ## will add constraints
lp_primal <- make.lp(0, 2)

# I will run the code below in a hidden chunk to avoid the output
# lp.control(lp_primal, sense = "max") 

```


```{r echo = FALSE, include = FALSE}
# Run as maximum
lp.control(lp_primal, sense = "max")

```

</div>

<div style="float: right; width: 50%;">

```{r}
# Set up model
set.objfn(lp_primal, c(40, 25))
add.constraint(lp_primal, c(30, 20), "<=", 1200)
add.constraint(lp_primal, c(10, 10),"<=", 500)
add.constraint(lp_primal, c(1, 0), "<=", 30)

# Add names to problem
dimnames(lp_primal) <- list(c("Labor", "Machines", "B-demand"),
                            c("Boots", "Shoes"))

# See the model we have set up
lp_primal

```

</div>


Solve model and look at solution
=============================================

<div style="float: left; width: 50%;">

```{r}
# Solve model
solve(lp_primal)

# Output, gives i) obj, ii) 3 constraint values and iii) 2 variable values
get.primal.solution(lp_primal) 

```

+ What do the results indicate  

  - 0 indicates solves  
  - 1575, objective function    
  - 1200, amount of labor used      
  - 450, amount of machine hours used  
  - 30, amount of boots produced, related to boots demand constraint  
  - 30, boots produced as a variable  
  - 15, shoes produced as a variable  

</div>

<div style="float: right; width: 50%;">

```{r}
# For two dimmension problems can plot the problem
plot(lp_primal)

```

</div>


View dual solution
=============================================

```{r}
# Dual solution
get.dual.solution(lp_primal)

```

+ What do the results indicate  

  - 1.00 indicates the dual is feasible  
  - 1.25, if increase L by 1 unit, profit increase by $1.25    
  - 0.00, if increase M by 1 unit, profit increase by $0  
  - 2.50, if increase D by 1 unit, profit increase by $2.50  
  - 0.00 and 0.00 indicate no "slack" in the variables  
    
+ **How can profit increase by $1.25 or $2.50 if boots are $40 and shoes $25?**  


Set up the dual problem so can compare results
=============================================

<div style="float: left; width: 50%;">

+ Program the same problem from dual perspective  

  - The dual has 3 variables  
  - Dual has 2 constraints with inequalities reversed  

```{r}
# Define the model with 0 constraints, is a minimum so not need lp.control()
lp_dual <- make.lp(0, 3)

# Add the constraints 
set.objfn(lp_dual, c(1200, 500, 30))
add.constraint(lp_dual, c(30, 10, 1), ">=", 40)
add.constraint(lp_dual, c(20, 10, 0),">=", 25)

dimnames(lp_dual) <- list(c("Boots", "Shoes"),
                            c("Labor", "Machines", "B-demand"))

```

</div>

<div style="float: right; width: 50%;">

```{r}
# See the model we have set up
lp_dual

```

</div>


Solve "the dual" model and look at solution
=============================================

<div style="float: left; width: 50%;">

+ To be clear  

  - This is a code of the dual to the original problem  
  - The dual of this is the primal of the model we first set up  
  - Mathematcially, it is arbitrary which is the primal and which is the dual  
  - Conceptually, set the primal as the model that makes the most sense for the application  
    
```{r}
# Solve model
solve(lp_dual)

# Output, gives i) obj, ii) 2 constraints values, and iii) 3 variable values
get.primal.solution(lp_dual) 

```

</div>

<div style="float: right; width: 50%;">

+ What do the results indicate  

  - 0 indicates solves  
  - 1575, objective function    
  - 45.00, profit of boots      
  - 25.00, profit of shoes  
  - 1.25, value of additional labor hours   
  - 0, value of additional machine hours   
  - 2.50, value of additional boot demand     

</div>


Solve "the dual" model and look at solution
=============================================


<div style="float: left; width: 50%;">

```{r}
# Output, gives i) obj, ii) 2 constraint values and iii) 3 variable values
get.dual.solution(lp_dual)

```

</div>

<div style="float: right; width: 50%;">

+ What do the results indicate  

  - 1.00 indicates the dual is feasible  
  - 30, if net profit of B increase by 1 dollar, profit increase by $30    
  - 15, if net profit of S increase by 1 dollar, profit increase by $15    
  - 0, all of L is used    
  - 50, M not used  
  - 0, all of D is used   

</div>


Sensitivity analysis
=============================================

+ Over what variable and constraint ranges will the solution hold?  

```{r}
# Get sensitivity analysis of primal
get.sensitivity.obj(lp_primal)
get.sensitivity.rhs(lp_primal)

```

+ Hard to tell what output is so will make a table  

Sensitivity analysis table
=============================================

<div style="float: left; width: 50%;">

```{r}
# Build a table to make understanding the results easier
obj_fn <- c(40, 25)
ps <- get.primal.solution(lp_primal) 
obj_sa <- get.sensitivity.obj(lp_primal)
rhs_sa <- get.sensitivity.rhs(lp_primal)

# Table dimmensions
n <- length(get.variables(lp_primal))
m <- length(get.constr.type(lp_primal))

# Make output and table
ov <- paste0("Objective function value = ", ps[1])
sa_tab <- rbind(ps[2:(n + m + 1)], 
                c(round(rhs_sa$duals[1:m], 2), obj_fn),
                round(c(rhs_sa$dualsfrom[1:m],obj_sa$objfrom), 2),
                round(c(rhs_sa$dualstill[1:m],obj_sa$objtill), 2)) 
colnames(sa_tab) <- c(rownames(lp_primal), colnames(lp_primal))
rownames(sa_tab) <- c("Solution", "Duals/Coef", "Sens From", "Sens Till")      

# Substitute "inf" and "-inf" so tables are easier to read (coded by two classmates!)
sa_tab <- ifelse(sa_tab == -1.000e+30, "-inf", sa_tab)
sa_tab <- ifelse(sa_tab == 1.000e+30, "inf", sa_tab)

```

</div>

<div style="float: right; width: 50%;">

```{r}
# Objective value and sensitivity analysis table
ov
as.table(sa_tab)

```

</div>


Sensitivity analysis
=============================================

+ We can format the table to be easier to read

```{r}
# Beautify the table using knitr and kableExtra packages, and add %>% 
`%>%` <- magrittr::`%>%`
library(knitr)
library(kableExtra)

# Print the table
kable(sa_tab, format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("striped", "bordered")) %>% 
  add_footnote(label = ov, notation = "none")

```

+ What if we add 1 more unit of Boot demand as discussed earlier?


Sensitivity analysis
=============================================

<div style="float: left; width: 50%;">

```{r echo = FALSE, include = FALSE}
# Set up model but eliminate verbose output
lp_primal_1 <- make.lp(0, 2)
lp.control(lp_primal_1, sense = "max")

# Continue model set up
set.objfn(lp_primal_1, c(40, 25))
add.constraint(lp_primal_1, c(30, 20), "<=", 1200)
add.constraint(lp_primal_1, c(10, 10),"<=", 500)
add.constraint(lp_primal_1, c(1, 0), "<=", 31)

# Add names to problem
dimnames(lp_primal_1) <- list(c("Labor", "Machines", "B-demand"), c("Boots", "Shoes"))

# Solve model
solve(lp_primal_1)

# Build a table to make understanding the results easier
obj_fn <- c(40, 25)
ps <- get.primal.solution(lp_primal_1) 
obj_sa <- get.sensitivity.obj(lp_primal_1)
rhs_sa <- get.sensitivity.rhs(lp_primal_1)

# Table dimmensions
n <- length(get.variables(lp_primal_1))
m <- length(get.constr.type(lp_primal_1))

# Make output and table
ov <- paste0("Objective function value = ", ps[1])
sa_tab <- rbind(ps[2:(n + m + 1)], 
                c(round(rhs_sa$duals[1:m], 2), obj_fn),
                round(c(rhs_sa$dualsfrom[1:m],obj_sa$objfrom), 2),
                round(c(rhs_sa$dualstill[1:m],obj_sa$objtill), 2)) 
colnames(sa_tab) <- c(rownames(lp_primal_1), colnames(lp_primal_1))
rownames(sa_tab) <- c("Solution", "Duals/Coef", "Sens From", "Sens Till")      

# Substitute "inf" and "-inf" so tables are easier to read (coded by two classmates!)
sa_tab <- ifelse(sa_tab == -1.000e+30, "-inf", sa_tab)
sa_tab <- ifelse(sa_tab == 1.000e+30, "inf", sa_tab)

```


```{r echo = FALSE}
# Print the table
kable(sa_tab, format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("striped", "bordered")) %>% 
  add_footnote(label = ov, notation = "none")

```

</div>

<div style="float: right; width: 50%;">

```{r}
# For two dimmension problems can plot the problem
plot(lp_primal_1)

```

</div>


