---
title: "Week 1: Constrained Optimization and Linear Programming"
author: "Gareth Green"
output: slidy_presentation
---

# Code basics

+ Recall, all code should have the following:  

  - Header with title, purpose, date and author  
  - Code documentation indicating each step  
  - Project tables and graphs should be professonial  
    - Exploritory tables and graphs can be rough  

```{r}
# Course: 5260 Math models for decision making
# Title: Week 1 - Constrained Optimization and Linear Programming
# Purpose: Basic demonstration of lpSolve and lpSolveAPI
# Date: March 1, 2020
# Author: Gareth Green

```


# Clear environment and load libraries

```{r}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

```


```{r}
# Load lpSolve package to demonstrate simple LP
library(lpSolve)

```


# Set up boots and shoes problem

+ B and S are decision variables  

+ Set objective function and constraints  

**Maximize over B and S**  
  40B + 25S  

**Subject to**  
  30B + 20S <= 1200  
  10B + 10S <=  500  
    B       <=   30  


# Code for basic LP problem using lpSolve

```{r}
# Define parameters of the lp command in lpSolve

# Name the objective function values
obj <- c(40, 25)

# Set up a matrix of constraints
con <- matrix (c(30 , 20, 
                 10, 10, 
                 1, 0), 
                 ncol = 2, byrow = TRUE)

# Set equation inequalities for the constraints
eq <- c("<=", "<=", "<=")

# Set the resource constraint values
rhs <- c(1200, 500, 30)

# Solve model
sol <- lp("max", obj, con, eq, rhs, compute.sens = TRUE)

```


# View the constraints, objective value and solution

```{r}
# View constraints
sol$constraints

```


```{r}
# lp output: Objective function value
sol$objval

```

```{r}
# lp output: Decision variables values
sol$solution

```

# Can solve the same problem using minimization  

+ **A max can be solved as a min, and vise versa**  

+ Set objective values as negative  

+ Set call in lp() to "min"  

  - All else is the same  
    
```{r}
# Define parameters of the lp command in lpSolve
obj <- c(-40, -25)

# Solve model
sol <- lp("min", obj, con, eq, rhs, compute.sens = TRUE)

# Use the negative of the objective function value
sol$objval
sol$solution

```


# Demo **lpSolveAPI** package  

+ lpSolveAPI has more functionality  

  - More to set up model  
  - But easier to add additional constraints  
  - Better sensitity analysis results  

```{r warning = FALSE}
# Load lpSolveAPI
library(lpSolveAPI)

# Define parameters of the lp

# This indicates 3 constraints and 2 variables
lpAPI <- make.lp(3, 2)

# Using lp.control() to set maximization problem, also gives additional printout of settings
lp.control(lpAPI, sense="max")

```


# Set up the model  

After set make.lp() can set up the model  

```{r}
# Objective function
set.objfn(lpAPI, c(40, 25))

# Set up constraints
set.row(lpAPI, 1, c(30, 20))
set.row(lpAPI, 2, c(10, 10))
set.row(lpAPI, 3, c(1, 0))

# Set inequalities for the constraints
set.constr.type(lpAPI, c("<=", "<=", "<="))

# Set the resource constraint values
set.rhs(lpAPI, c(1200, 500, 30))

# See the model we have set up
lpAPI

```


# Solve model

+ 0 Returned indicates model was successfully solved 

+ See ?solve.lpExtPtr for solve codes

```{r}
# Solve the model, solve() is a lpSolveAPI function
solve(lpAPI)

# Output: Objective function value
get.objective(lpAPI)

# Output: Decision variables and constraint values
get.variables(lpAPI)
get.constraints(lpAPI) 

```

# Can add aditional constraints

+ One benefit of lpSolveAPI is you can easily add constraints  

  - This make constructing the model easier  

+ Use the same model, but add a constraint  

+ Add a ratio constraint  

  - 3B <= S: S has to be more than 3 times B  
  - Maybe a materials or shipping constraint  
  - Mathematically this is 3B - S <= 0  
  - Cannot have variable on the right hand side of constraint  

```{r}
# Add a new type of constraint 3B <= S
add.constraint(lpAPI, c(3, -1), "<=", 0)

```

+ **Do you expect profit to be more, less or same**  

  - What does change in profit depend on?  

# New model

+ Additional constraints will reduce profitability if they are binding

  - If binding, the constraint has made the feasibility region smaller
  - If not binding, no impact

```{r}
# Solve model
solve(lpAPI)

# Output: Objective function value
get.objective(lpAPI)

# Output: Decision variables values
get.variables(lpAPI)
get.constraints(lpAPI) 

```


# Wrap up

+ lpSolve and lpSolveAPI are farily straight forward packages for linear programing

  - As problems grow the coding of constraints can become challenging
  - Provide information on solution

+ R is not designed to be linear programing software

  - Can also use Python, but not much better
  - If actually doing large LP problems will likely be using LP specific software
  - Can also use Excel for small problems
  
+ R helps you learn the basic structure to understand constrained optimization

  - Goal of this class is to educate you about constrained optimization
    - important modeling framework used extensively modeling
