---
title: "Week 8"
subtitle: "Queueing Theory, Ch 17"
author: "Gareth Green"
output: slidy_presentation
---

```{r echo = FALSE}
# Course: 5260 Math models for decision making
# Title: Week 8 queueing theory, Ch 17
# Purpose: Demonstrate applications of queueing models 
# Date: March 15, 2020
# Author: Gareth Green

```

```{r echo = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

```

```{r echo = FALSE}
# Load "queueing" package
library(queueing)
`%>%` <- magrittr::`%>%`
library(knitr)
library(kableExtra)
library(ggplot2)

```
  

Typical queueing model questions
=============================================

<div style="float: left; width: 50%;">

+ We use a number of formulas from the text derived from steady state  

    - Derivations are long and detailed so not duplicate what is covered in the text  

+ Answer several questions  

    1. Probability of _n_ people in queue system or queue  
    2. Probability wait in queue system or queue $\gt t$, pg 752-3
       - where $t$ is a specified time period to wait the business feels is reasonable
    3. Mean number of customers in queue system or queue
    4. Mean wait time in queue system or queue 

</div>

<div style="float: right; width: 50%;">

+ The utilization factor is important $\rho = \lambda /s \mu$, pg 735

  - $\rho$ is the the expected proportion servers are busy  
    - Represents the the system's service capacity $s\mu$, (servers * customer time/service) 
    - that is being _utilized_ by arriving customers at rate $\lambda$, customers needing service

+ Related but determined separately is the cost of waiting and cost of additional servers

  - these determine the economic trade-off of hiring more servers to reduce wait time

</div>


Problem set up
=============================================

<div style="float: left; width: 40%;">

+ Example #1, M/M/1/inf/inf/FCFS, aka M/M/1

  - Two customers arrive per hour  
  - It takes 15 min (i.e. 4 per hour) to complete service. 

+ We will use two approaches:

  - first we will make calculations based on formulas derived in the text
    - this illustrates the power of using the **steady state** assumption
  - second we will use the `queueing` package to examine questions in more detail
    - will also allow us to answer questions in more complicated models

</div>

<div style="float: right; width: 60%;">

+ Begin with estimating the probability of customers and wait times given one server
  - note _s_ = 1 is the special case of an M/M/1 where $\rho = \lambda/\mu = 2/4 = 0.5$
  - you should expect your server to be busy (i.e. utilized) half of the time
  - must have $\rho < 1$ otherwise the queue will grow to infinity
    - i.e. more people than capacity

```{r}
# Set parameters
lam <- 2
mu <- 4
s <- 1 # One server case
rho <- lam/(s*mu)

# Probability of 0 people, see pg 751
p0 <- 1 - rho 

# Let's look for up to 5 people
n <- 5

# Build a table of results, go to n + 1 so can see Pn when there are 0 customers
tab_1 <- as.data.frame(matrix(0, n + 1, 7))

```

</div>


Code the calculations
=============================================

<div style="float: left; width: 50%;">

+ We're not using any functions, just making calculations based on formula

```{r}
# Set a loop to make steady state calculations for each level of i customers in system
for (i in 0:n){
  # Probability of i people in queue system, see pg 751
  Pn <- p0*rho^i 

  # Mean number of customers in queue system, see pg 751-2
  L <- lam/(mu - lam)  
  
  # Mean wait time in queue system, by Little's Law pg 736
  W <- L/lam 
  
  #Mean wait time in queue waiting for service, by Little's Law pg 736
  Wq <- W - 1/mu 
  
  # Mean number of customers in queue, by Little's Law pg 736
  Lq <- lam * Wq 
  
  # Combine in a table
  tab_1[i + 1, ] <- round(c(i, Pn, s, L, Lq, W*60, Wq*60), 3) 
}

# Put header on table and name the header in the same line of code
colnames(tab_1) <- c_name <- c("Customers", "Pn", "s", "L", "Lq", "W (minutes)", "Wq (minutes)")

```

</div>

<div style="float: right; width: 50%;">

```{r echo = FALSE}
# Show results, note that L, Lq, W, Wq remain constant
tab_1 %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "bordered"))

```

+ Recall what each of these variables represents? 

  - Pn is the probability of n people in the system

+ Why are the other variables constant?

  - because they are expected values in a steady state

<br>

</div>


Can make similar calculations using the `queueing` package
=============================================

<div style="float: left; width: 50%;">

+ `queueing` is an impressive package

  - check the [documentation](https://www.rdocumentation.org/packages/queueing/versions/0.2.12)

+ What was the main reason we were able to make easy calculations above??

+ Now use the queueing package to make the same calculations  

  - We can make own calculations easily for M/M/1 model  
  - Gets much more complicated for other models so use `queueing` package  

+ For example, if $s = 1$, then $P_0 = 1 - \rho$

  - but for $s > 1$, $P_0$ is a massive equation, see top pg 754

</div>

<div style="float: right; width: 50%;">

+ We can pull the output from each model run for different number of customers

```{r }
# create memory for a table of results, expand table
n <- 10
tab_q1 <- as.data.frame(matrix(0, n+1, 7))

# Set a loop to make steady state calculations for each level of i customers in system
for (i in 0:n){
  # Using "queueing" package
  q_mod1 <- NewInput.MM1(lam, mu, i)
  m1 <- QueueingModel.i_MM1(q_mod1)

  # Place results as a row in a table
    # Multiply wait by 60 minutes in an hour to convert time from hours to minutes
  tab_q1[i+1,] <- round(c(i, m1$Pn[i+1], s, m1$L, m1$Lq, m1$W*60, m1$Wq*60), 3) 
}

# Add column names
colnames(tab_q1) <- c_name

```

</div>


Model results
=============================================

<div style="float: left; width: 50%;">

```{r echo = FALSE}
# Show results, note that L, Lq, W, Wq remain constant
tab_q1 %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "bordered"))

```

</div>

<div style="float: right; width: 50%;">

```{r echo = FALSE}
# Visualize the relationship between Pn
tab_q1 %>%
  ggplot(mapping = aes(x = Customers, y = Pn)) +
  geom_line() +
  labs(title = "The probability of customers in the system is decreasing", y = "Probability of customers") +
  scale_x_continuous("Potential customers", breaks = seq(0, 10, by = 2)) +
  theme_classic()

```

</div>

Other output from queueing
=============================================

<div style="float: left; width: 45%;">

+ Important to pay attention to the definitions because there are many possibilities

+ Can use queuing package to calculate other values too

  - Probability all servers are busy

```{r}
# Probability all servers are busy
B_erlang(s, lam/mu)

```

+ Probability to wait in queue because all servers are busy

```{r}
# Probability to wait in queue because all servers are busy
C_erlang(s, lam/mu)

```

+ C_erlang is higher because there is servers being busy and other customers in the queue

+ These are different than $P_n$ which is probability of $n$ customers in the system

</div>

<div style="float: right; width: 55%;">

+ Another way to get output from the model

```{r}
# Use the Report() function
Report(m1)

```

</div>


Example #2, M/M/3/inf/inf/FCFS aka M/M/3
=============================================

<div style="float: left; width: 50%;">

+ A call center which has an arrival rate of 15 calls per hour  

  - Each query take approximately 10 minutes ($\mu = 6$ per hour) to be resolved  
  - Suppose there are 3 servers

+ The utilization rate is $\rho = \lambda/(s*\mu) = 15/(3*6) = 0.83$

+ **What are the system outcomes??** 

</div>

<div style="float: right; width: 50%;">

```{r}
# Set up problem with criteria and 8 customers
n <- 8; lam = 15; mu = 6; s = 3 

# Build a table of results
tab_q2 <- as.data.frame(matrix(0, n+1, 7))

# Loop to build table of results
for (i in 0:n){
  
  # Using "queueing" package
  q_mod2 <- NewInput.MMC(lambda = lam, mu = mu, c = s, n = n)
  m2 <- QueueingModel.i_MMC(q_mod2)
  
  # Place results as a row in a table
    # Multiply wait by 60 minutes in an hour to convert time from hours to minutes
  tab_q2[i+1,] <- round(c(i, m2$Pn[i + 1], m2$Inputs$c, m2$L, m2$Lq, m2$W*60, m2$Wq*60), 3) 
}

# Add column names
colnames(tab_q2) <- c_name

```

</div>


Model output
=============================================

<div style="float: left; width: 50%;">

+ Average 6 people in the system and 3.5 in the queue

  - Average 24 minutes in the system with 14 minutes waiting in the queue
  - 14  minutes is a long time to wait for a 10 minute service

```{r echo = FALSE}
# Show results, note that L, Lq, W, Wq remain constant
tab_q2[1:6, ] %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "bordered")) %>% 
  add_footnote(paste("Probability all servers are busy is", round(B_erlang(s, lam/mu), 3))) %>% 
  add_footnote(paste("Probability customers will wait in the queue", round(C_erlang(s, lam/mu), 3)))

```

</div>

<div style="float: right; width: 50%;">

```{r echo = FALSE}
# Visualize the relationship between Pn
tab_q2 %>%
  ggplot(mapping = aes(x = Customers, y = Pn)) +
  geom_line() +
  labs(title = "Most likely to have two customers in the system", 
       subtitle = "The probability of 0 and 8 customers is about equal",
       y = "Probability of customers") +
  scale_x_continuous("Potential customers", breaks = seq(0, n, by = 2)) +
  theme_classic()

```

</div>


In M/M/C, how do variables change with Servers
=============================================

<div style="float: left; width: 50%;">

+ We saw that with 3 servers there is most likely going to be:

  1. Six customers in the system
  2. Twenty-four minutes in the system
  3. Fourteen minutes in the queue

+ But these are long wait times in the queue

  - How many servers needed to reduce queue time to a reasonable amount
  
</div>

<div style="float: right; width: 50%;">

+ Here we will loop over the number of servers

```{r}
# Number of cutomers, Initial servers
n <- 8; int <- 3

# Build a table of results
tab_q3 <- as.data.frame(matrix(0, n+1-int, 7))

# Loop to build table of results
for (s in int:n){

  # Using "queueing" package
  q_mod3 <- NewInput.MMC(lambda = lam, mu = mu, c = s, n = n)
  m3 <- QueueingModel.i_MMC(q_mod3)
  
  # Place results as a row in a table
    # Multiply wait by 60 minutes in an hour to convert time from hours to minutes
  tab_q3[s+1-int,] <- round(c(s, m3$L, m3$Lq, m3$W*60, m3$Wq*60, B_erlang(s, lam/mu), C_erlang(s, lam/mu)), 3) 
}

# Add column names
colnames(tab_q3) <- c("Servers", "L", "Lq", "W (minutes)", "Wq (minutes)", "B_erlang", "C_erlang")

```

</div>


Model results
=============================================

<div style="float: left; width: 55%;">

+ Wow, one more server reduces the wait time to just over 2 minutes

  - and reduce the number of people waiting in the queue by 3
  - getting 5 servers (2 more) reduces wait time to just over 30 seconds

```{r echo = FALSE}
# Show results, note that L, Lq, W, Wq remain constant
tab_q3 %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "bordered"))

```

</div>

<div style="float: left; width: 45%;">

+ Easy to see why companies want to automate queues

  - can reduce queue time dramatically with low cost of automation
  - at least keep customers engaged as part of the queue process at low cost

```{r echo = FALSE, fig.height=4, fig.width=6}
# Visualize the relationship between Pn
tab_q3 %>%
  ggplot(mapping = aes(x = Servers, y = `Wq (minutes)`)) +
  geom_line() +
  labs(title = "Big reduction in queue time going from 3 to 4 servers") +
  scale_x_continuous("Servers", limits = c(2, 8), breaks = seq(2, n, by = 2)) +  
  scale_y_continuous("Minutes in the queue", limits = c(0, 15)) +
  theme_classic()

```

</div>


Work through a complete question
=============================================

<div style="float: left; width: 90%;">

Several students had a start up idea of helping people with their software by using a software package that allows them to easily control someones computer remotely with their permission. Softie.com's goal is to deliver personal service over the internet. They believe there is a niche market for people that want to see and talk to actual people rather than bots. Softie.com is a student team trying to figure out how to launch and has the following parameters:

+ 2 server call center (to get started)
+ 9 inquiries per hour (based on an estimate from industry data)
+ Each query take approximately 12 minutes (5 per hour) to be resolved

+ What is the utilization rate?

  >- The utilization rate indicates the proportion a system's servers are in use
  >- $\rho = \lambda/(s*\mu) = 9/(2*5) = 0.9$
  >- They are at a high utilization so they are likely to have some long wait times in the queue
  >- It may be difficult to make it in a niche market with a goal of quality and long wait times
  >- **Set up the problem in R**

<br> 

</div>

Set up the problem
=============================================

<div style="float: left; width: 50%;">

+ We will use the `queueing` package since $s > 2$

```{r}
# Set up problem with criteria and 10 customers
lam = 9; mu = 5; s = 2; n <- 20;  

# Using "queueing" package
soft_mod <- NewInput.MMC(lambda = lam, mu = mu, c = s, n = n)
sm <- QueueingModel.i_MMC(soft_mod)

```

+ Though the probabilities are low, they drop off slowly

  - The graph is over 20 customers
  - This indicates long queue wait times

</div>

<div style="float: right; width: 50%;">

```{r echo = FALSE}
# Graph wait time in the servers vs wait time in the queue 
ggplot(mapping = aes(x = seq(0:n-1), y = sm$Pn)) +
  geom_line() +
  theme_classic() +
  theme(axis.text.x=element_text(hjust = -3), # This realigns the data with the correct customer
        axis.ticks = element_blank()) +
  scale_x_continuous("Customers", breaks = seq(0, n, by = 2), limits = c(0, n+1 )) +
  scale_y_continuous("Probability of Customers", limits = c(0, max(sm$Pn)))

```

</div>


Get full model results
=============================================

<div style="float: left; width: 50%;">

+ Any way you look at it these numbers are bad!

  - W = `r round(sm$W * 60, 0)` minutes
  - Wq = `r round(sm$Wq * 60, 0)` minutes

+ People won't wait that long for service even if it is amazing

+ The problem with hiring more servers is that the utilization rate drops 

<div style="float: left; width: 45%;">

  - Servers sit idle due to randomness of arriving customers
  - Idle servers is expensive
  - you can see $\rho$ drops off slowly

</div>

```{r echo = FALSE}
# Make table showing utilization rate
data.frame(Servers = seq(1:7)) %>%  
  dplyr::transmute(Servers = Servers + 1) %>% 
  dplyr::mutate(rho = round(lam/(Servers * mu), 2)) %>%
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "bordered"), full_width = FALSE)  

```

</div>

<div style="float: right; width: 50%;">

```{r echo = FALSE}
# Set up problem with criteria and 10 customers
Report(sm)

```

</div>


What about more servers??
=============================================

<div style="float: left; width: 60%;">

+ The student team is reluctant to hire more servers

  - don't want servers sitting idle
  - they hatch an idea that idle servers can provide additional services

+ **What is the probability currently with two servers that one could work on an optional service?**

  >- That would be $\sum_0^1 P_i = P_0 + P_1$ = `r round(sum(sm$Pn[1:2]), 3)`
  >- That is pretty low, how much do think it would increase with another server?

+ **What is the probability of time to work on other services if hire a third server?**

```{r}
# Set up problem with criteria and 20 customers
lam = 9; mu = 5; s = 3; n <- 20;  

# Using "queueing" package
soft_mod <- NewInput.MMC(lambda = lam, mu = mu, c = s, n = n)
sm <- QueueingModel.i_MMC(soft_mod)
  
```

  >- That would be $\sum_0^2 P_i = P_0 + P_1 + P_2$ = `r round(sum(sm$Pn[1:3]), 3)`

<br>

</div>

<div style="float: right; width: 40%;">

```{r echo = FALSE}
# Graph customers vs probability of customers in the system
ggplot(mapping = aes(x = seq(0:n-1), y = sm$Pn)) +
  geom_line() +
  theme_classic() +
  theme(axis.text.x=element_text(hjust = -3), # This realigns the data with the correct customer
        axis.ticks = element_blank()) +
  scale_x_continuous("Customers", breaks = seq(0, n, by = 2), limits = c(0, n+1 )) +
  scale_y_continuous("Probability of Customers", limits = c(0, max(sm$Pn + 0.05)))

```

</div>


Get full model results
=============================================

<div style="float: left; width: 50%;">

+ This is substantially better, servers could have time for an additional product!

  - W = `r round(sm$W * 60, 0)` minutes
  - Wq = `r round(sm$Wq * 60, 0)` minutes

+ What are other results?

  - you can see the Pn drops off quickly

+ Though the physical numbers look good, what does this trade-off come down to? 

  >- **The cost of servers vs the cost of waiting**

</div>

<div style="float: right; width: 50%;">

```{r echo = FALSE}
# Set up problem with criteria and 10 customers
Report(sm)

```

</div>


How does hiring servers impact time in the queue
=============================================

<div style="float: left; width: 50%;">

+ Suppose Softie determines they can charge more if they can get the queue time to less than 30 seconds

  >- How many servers will they need to hire?

+ Here we will loop over the number of servers

```{r}
# Set up problem with initial criteria and 20 customers
lam = 9; mu = 5; s = 2; n <- 8; int <- 2

# Build a table of results
tab_smc <- as.data.frame(matrix(0, n+1-int, 7))

# Loop to build table of results
for (s in int:n){

  # Using "queueing" package
  sm_c <- NewInput.MMC(lambda = lam, mu = mu, c = s, n = n)
  smc <- QueueingModel.i_MMC(sm_c)
  
  # Place results as a row in a table
    # Multiply wait by 60 minutes in an hour to convert time from hours to minutes
  tab_smc[s+1-int,] <- round(c(s, smc$L, smc$Lq, smc$W*60, smc$Wq*60, B_erlang(s, lam/mu), C_erlang(s, lam/mu)), 3) 
}

# Add column names
colnames(tab_smc) <- c("Servers", "L", "Lq", "W (minutes)", "Wq (minutes)", "B_erlang", "C_erlang")

```

</div>

<div style="float: right; width: 50%;">

```{r echo = FALSE}
# Visualize the relationship between Pn
tab_smc %>%
  ggplot(mapping = aes(x = Servers, y = `Wq (minutes)`)) +
  geom_line() +
  labs(title = "Big reduction in queue time going from 2 to 3 servers") +
  scale_x_continuous("Servers") +  
  scale_y_continuous("Minutes in the queue") +
  theme_classic()

```

</div>


Numerical results show...
=============================================

<div style="float: left; width: 50%;">

```{r echo = FALSE}
# Show results, note that L, Lq, W, Wq remain constant
tab_smc %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "bordered"))

```

</div>

<div style="float: right; width: 45%;">

+ Going from 2 to 3 is huge, but probably need to go to 4 servers

  - 3 servers still has a 3.5 minute wait
  - 4 servers has less than a 45 second wait

+ What is the probability that at least one of the call servers will be available to provide additional services?

  >- recall B_erlang is the probability all servers are busy
  >- C_erlang is the probability will have to wait in line
  >- numbers for 4 serves shows small probability of all servers being busy and customers waiting in line

</div>


What if scenarios
=============================================

<div style="float: left; width: 50%;">

Assume Softie.com purchased 2 more servers for a total of 4. But now Softie.com's incoming business doubles. How many servers will they need more to maintain a less-than one-minute queue time?

```{r}
# Set up problem with initial criteria and 20 customers
lam = 9 * 2; mu = 5 ; s = 2; n <- 10; int <- 4

# Build a table of results
tab_smc <- as.data.frame(matrix(0, n+1-int, 7))

# Loop to build table of results
for (s in int:n){

  # Using "queueing" package
  sm_c <- NewInput.MMC(lambda = lam, mu = mu, c = s, n = n)
  smc <- QueueingModel.i_MMC(sm_c)
  
  # Place results as a row in a table
    # Multiply wait by 60 minutes in an hour to convert time from hours to minutes
  tab_smc[s+1-int,] <- round(c(s, smc$L, smc$Lq, smc$W*60, smc$Wq*60, B_erlang(s, lam/mu), C_erlang(s, lam/mu)), 3) 
}

# Add column names
colnames(tab_smc) <- c("Servers", "L", "Lq", "W (minutes)", "Wq (minutes)", "B_erlang", "C_erlang")

```

</div>

<div style="float: right; width: 50%;">

```{r echo = FALSE}
# Visualize the relationship between Pn
tab_smc %>%
  ggplot(mapping = aes(x = Servers, y = `Wq (minutes)`)) +
  geom_line() +
  labs(title = "Big reduction in queue time going from 2 to 3 servers") +
  scale_x_continuous("Servers") +  
  scale_y_continuous("Minutes in the queue") +
  theme_classic()

```

</div>


Numerical results show...
=============================================

<div style="float: left; width: 50%;">

```{r echo = FALSE}
# Show results, note that L, Lq, W, Wq remain constant
tab_smc %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "bordered"))

```

</div>

<div style="float: right; width: 45%;">

+ Going from 4 to  5 is huge, but need to go to 6 servers to be less than 1 minute wait

  - 5 servers still has a 3.5 minute wait
    - similar as going from 2 to 3 servers before
  - 6 servers has less than a 1 minute wait
    - before going from 3 to 4 had a 45 second wait
    
+ Their business doubles but they only need to increase their servers by 50% to maintain a similar level of wait time in the queue

  - of course this is mean time
  - if we did a simulation we would see that some wait times are substantially larger and some substantially smaller
    - we will look at this in the next module
    - but Poisson is asymmetric to to you expect there to be more long waits or short waits?
    - this is a case where the mean may not be a good measure

<br>

</div>


What if have cost data
=============================================

<div style="float: right; width: 90%;">

+ What if we knew the cost of lost customers and servers??  

  - Still not an optimization problem like in earlier chapaters  
  - But we can find the optimal number of servers  
    
+ **Problem set up**, same as before  

  - A call center which has an arrival rate of 18 queries per hour  
  - Each query take approximately 12 minutes (5 per hour) to be resolved   
  - What is the cost minimizing number of servers if:  
    - A server costs $15/hour  
    - The firm loses $40 per lost customer
    - Cusomers will leave if the wait is more than 30 seconds

</div>


Code the problem set up
=============================================

<div style="float: left; width: 50%;">

+ Similar to other queueing theory problems, but add costs

```{r}
# Set up problem with initial criteria and 10 customers
lam = 18; mu = 5 ; s = 2; n <- 10

# Minimum number of servers so queue not grow to infinity, ceiling rounds up to next integer
int <- ceiling(lam/mu)

# Profit per customer
pi <- 40

# Cost per server per hour
cost <- 15

# Build a table of results
tab_smc <- as.data.frame(matrix(0, n+1-int, 8))

```

</div>

<div style="float: right; width: 50%;">

+ Note the time threshold of 30 seconds (0.5 minutes) for losing customers

```{r}
# Loop to build table of results
for (s in int:n){

  # Using "queueing" package
  sm_c <- NewInput.MMC(lambda = lam, mu = mu, c = s, n = n)
  smc <- QueueingModel.i_MMC(sm_c)

  # Cost of lost customers
  lp <- ifelse(smc$Wq*60 > 0.5, smc$Lq * pi, 0)
  # Cost of servers
  sc <- s*cost  
  # Total cost
  tc <- sc + lp
  
  # Place results as a row in a table
    # Multiply wait by 60 minutes in an hour to convert time from hours to minutes
  tab_smc[s+1-int,] <- round(c(s, smc$L, smc$Lq, smc$W*60, smc$Wq*60, lp, sc, tc), 2) 
}

```

</div>


Results
=============================================

<div style="float: left; width: 45%;">

```{r echo = FALSE}
# Add column names
colnames(tab_smc) <- c("Servers", "L", "Lq", "W (minutes)", "Wq (minutes)", 
                       "Lost Profit", "Server Cost", "Total Cost")

# Show results, note that L, Lq, W, Wq remain constant
tab_smc %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "bordered"))

```

</div>

<div style="float: right; width: 45%;">

+ Look for minimum TC  

+ 6 servers gives the minimum cost of trading off between servers and lost customers  

  - Cost of lost customers decreases as cost of servers increases 
  - Note the cost of lost customers is still positive from 6 to 7 servers
    - inidicates it is optimal to lose some customers rather than hire additional servers
    - this is based on the probability of needing 7 servers vs the cost of hiring an aditional server

+ Does anyone remember why we are using this calculated approach?

</div>


Why use the numerical approach?
=============================================

<div style="float: left; width: 47%;">

+ This looks like a basic calculus problem

  - unconstrained minimization
  - just take the derivative of total cost with respect to *s*, right?
  - the function is complicated when $s > 1$
  
+ Total cost = lost profit  + server cost

  >- $TC =  L_q * \pi + s * cost$, which looks easy enough
  >- $L_q$ = ${P_0(\lambda/\mu)^s\rho}\over{s!(1 - \rho)^2}$ = ${P_0(\lambda/\mu)^s\lambda/(s\mu)}\over{s!(1 - \lambda/(s\mu))^2}$
  >- what is wrong with taking a derivative of this with respect to $s$?
  >- factorials are discontinuous so a derivative doesn't make sense
  >- Also, the equation for $P_0$ is ridiculous, top pg 754

</div>

<div style="float: right; width: 53%;">

```{r echo = FALSE}
# Graph the cost relationship
tab_smc %>% 
  ggplot(aes(x = Servers, y = `Lost Profit`)) +
  geom_line(aes(y = `Lost Profit`), color = "red")  +
  geom_line(aes(y = `Server Cost`), color = "blue")  +
  geom_line(aes(y = `Total Cost`)) +
  geom_hline(yintercept = min(tab_smc$`Total Cost`), linetype = "dotted") +
  labs(title = "Should hire 6 servers even though lose some customers", x = "Servers", y = "Cost of service") +
  theme_classic()

```

</div>

Wrap up
=============================================

<div style="float: left; width: 90%;">

+ Queues are everywhere so are important to analyze and understand

  - We barely scratched the surface of Queueing Theory
  - Many online resources and [calculators](http://erlang.chwyean.com/erlang/erlangB.html)
  - Work through the `queueing` package if you want to learn more

+ Can be very complex, we have only studied simple systems to understand the basics

  - Could easily spend a quarter just studying Queueing Theory

+ We focused on using the steady state assumption and Little's Law

  - This gave us good mean times to work with
  - Next week we will look at simulations to see how these variables vary around the mean values

</div>


