---
title: "Week 4"
subtitle: "Network examples in igraph"
author: "Gareth Green"
output: slidy_presentation
---
  
```{r echo = FALSE, message = FALSE}
# Course: BUAN 5260
# Title: Week 4-Network examples
# Purpose: Demo igraph package on example problems of Chapter 10
# Date: March 15, 2020
# Author: Gareth Green

# Clear packages 
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(
  paste("package:", names(sessionInfo()$otherPkgs), sep=""), 
  detach, character.only = TRUE, unload = TRUE)

# Clear environment
rm(list = ls(all = TRUE)) 

```

Set up network optimization in igraph package 
=============================================

+ We will demonstrate the igraph package for network optimization

  - though igraph can do network optimization it is not a constrained optimization package
  - however, there is a cross-over with classic and modern network optimization so want to demonstrate with igraph

+ We will use the Seervada Park Network to look at the following:

  - Shortest-path, 10.3 
  - Minimum spanning tree, Ch 10.4
  - Maximum flow, Ch 10.5

+ In many ways the igraph package is easier to use than lpsolveAPI

  - but you do not set up the problem in the classic sense
  - the objective function and constraints are set up in the graph of your network
  - do not get _sensitivity analysis_ results so provides less information
  - I have had students use it as a check on their work
  
```{r message = FALSE}
# Load packages
library(igraph)

```


Set up Seervada Park Network 
=============================================


```{r}
# Data matrix, could also import data file
seer <- matrix(nc = 4, byrow = TRUE, 
              c(1, 2, 2, 5, 
                1, 3, 5, 7, 
                1, 4, 4, 4, 
                2, 3, 2, 1, 
                2, 5, 7, 3, 
                3, 4, 1, 2,
                3, 5, 4, 4,
                3, 6, 3, 5,
                4, 6, 4, 4,
                5, 6, 1, 1,
                5, 7, 5, 9,
                6, 7, 7, 6)) 

# Make graph by setting 7 nodes/vertices, edges from nodes/vertices and distances from data matrix
g <- add_edges(make_empty_graph(n = 7, directed = FALSE), 
               t(seer[, 1:2]), weight=seer[,3]) %>%

  # Set vertice/node names 
  set_vertex_attr("name", value = c("O","A","B","C","D","E","T"))

# Set coordinates so looks similar to book diagram
  ## This is just for visualization, but could include it in data file
g$layout <- matrix(c(0,   25,  50,  40,  100, 90,  150, 
                     150, 175, 150, 125, 150, 125, 160), nc = 2)

# Lable edges with the distance or weight
E(g)$label <- seer[, 3]

```


Seervada Park Network
=============================================

```{r}
# Show network graph
plot(g)

```


Find shortest path(s), Ch 10.3
=============================================

```{r}
# Shortest path(s)
all_shortest_paths(g, "O", "T")$res[[1]]

# Show distance matrix between nodes/verticies
distances(g, "O", "T")

```


Find minimum spanning tree, Ch 10.4
=============================================

```{r}
# Minimum spanning tree
g_mst <- mst(g, weights = NULL)
g_mst
plot(g_mst)

```


Find maximum flow, Ch 10.5
=============================================

```{r}
# Maximum flow
g_max <- max_flow(g, "O", "T", capacity = seer[,4])
g_max$value
g_max$flow

# Plot max flow
gx <- g
E(gx)$label <- g_max$flow
plot(gx)

```

