---
title: "Week 9"
subtitle: "Queueing Simulation"
author: "Gareth Green"
output: slidy_presentation
---

```{r echo = FALSE}
# Course: 5260 Math models for decision making
# Title: Week 9 - Queuing Simulation
# Purpose: Demonstrate simulation using base R 
# Date: March 15, 2020
# Author: Gareth Green

```

```{r echo = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

```
  

```{r echo = FALSE, message = FALSE}
# Load library
library(queueing)
library(dplyr)
library(ggplot2)
library(patchwork)
`%>%` <- magrittr::`%>%`
library(knitr)
library(kableExtra)

```


Queueing simulation
=============================================

<div style="float: left; width: 50%;">

+ The benefit of running simulation is we can see the distribution of variables rather than just mean values

  - the mean values are good information
  - realistically customers will not arrive at mean rate and service will not take mean time

+ The simulated parameters will be arrival and service time

  - realize we could also simulate server outage and other important questions

+ The difficulty is that the `queueing` package does not support Monte Carlo simulation

  - as far as I can tell `queueing` only finds mean values for standard variables
  - valuable, but variation is important to understand

<br>

</div>

<div style="float: right; width: 50%;">

+ Begin by coding Monte Carlo Simulation of an M/M/1 in the `queueing` package

  - 3 people arrive per hour, lambda = 3
  - service takes 12 min, mu = 5
  - 1 server

+ Will build my own model to simulate a queueing system

  - the Will simulate 1000 runs of for 250 customers
    - 250 = 3 customers per hour * 12 hour day * 7 days per week
    - so simulate 1 week's worth of business 1000 times, about 19 years worth
  - will calculate mean times from queueing and compare to simulation
  - will examine distributions of key variables
    - L, Lq, W and Wq

</div>


Code the standard `queueing` set up
=============================================

<div style="float: left; width: 90%;">

+ First will find L, Lq, W and Wq using `queueing` package so we have a base line of comparison

  - when possible you want to have a comparison for results of your own model to verify accuracy of your results

```{r}
# Set up problem with initial criteria and 10 customers
lam <- 3; mu <- 5 ; s <- 1; 

# Using "queueing" package
q <- NewInput.MMC(lambda = lam, mu = mu, c = s)
q_mod <- QueueingModel.i_MMC(q)
# Place results as a row in a table
  # Multiply wait by 60 minutes in an hour to convert time from hours to minutes
q_tab <- data.frame(t(round(c(s, q_mod$L, q_mod$Lq, q_mod$W*60, q_mod$Wq*60), 2))) 

colnames(q_tab) <- c("Servers", "L", "Lq", "W (minutes)", "Wq (minutes)")

```


```{r echo = F}
# Show results, note that L, Lq, W, Wq remain constant
(q_tab %>% 
  kable(caption = "Queueing package mean steady state results") %>%
  kable_styling(bootstrap_options = c("striped", "bordered")))

```

</div>


Poisson and exponential distributions
=============================================

<div style="float: left; width: 50%;">

+ An interesting feature of the two distributions:

  - The Poisson shows the number of occurrences per time interval
  - The exponential shows the length of time between occurrences

+ To set up the simulation model I will model the length between intervals of arrivals and service time

  - I initially tried modeling the number of occurrences per time interval using the Poisson
  - it did not work because I could not model the time in the queue

+ It is not unusual when you build a model that you have to modify your approach to find the variables of interest

</div>

<div style="float: right; width: 50%;">

```{r}
# Number of random variables
n <- 1000

# Number of customers in each run, which is about 1 week worth of business if lam = 3 per hour
cust <- 250

# Number of runs, i.e. number of weeks
runs <- n

# Set seed to get consistent results while coding
set.seed(3) 

# Generate random variables
e_arr <- rexp(n, rate = lam)
e_ser <- rexp(n, rate = mu)

```

</div>


Coding the customer loop 
=============================================

<div style="float: left; width: 50%;">

+ The model will take two stages

  1. the model of customer arrival and service for a given number of customers
      - in this case 250, about a weeks worth of customers
  2. the loop to run the model many times
      - 1000 times in this case, about 19 years worth of weeks

+ When I first coded the model I used data.frame

  - once running I converted to matrix, waaaaaayyyyyyy faster
    - should have done that from the beginning
  - I could convert each loop to be a function but am short on time

+ So what is going on in this model!?

</div>

<div style="float: right; width: 50%;">

```{r eval = F}
# Build a table of results
tab <- matrix(0, cust, 7)

# Set initial time, wait time
arr_time <- 0; beg_time <- 0; end_time <- 0; wait_time <- 0

# Set up loop for every customer
for(i in 1:cust){
  # Minutes till next customer arrives to the system
  t_arr <- round(sample(e_arr, 1, replace = TRUE) * 60, 0)

  # Time at customer arrival
  arr_time <- arr_time + t_arr
  
  # Customer wait time
  wait_time <- ifelse(end_time > arr_time, end_time - arr_time, 0)
  
  # Set service begin time
  beg_time <- arr_time + wait_time 
  
  # Random service times for each customer
  t_ser <- round(sample(e_ser, 1, replace = TRUE) * 60, 0)
  
  # Set time service ends
  end_time <- beg_time + t_ser

  # Combine in a table
  tab[i, ] <- round(c(i, t_arr, arr_time, beg_time, t_ser, end_time, wait_time), 1) 
}

```

</div>


Coding the weeks loop 
=============================================

+ Need to run the 250 customers through 1000 weeks

  - this code does not have the customer loop in it yet

```{r eval = F}
# Initialize results table
total <- matrix(0, runs, 5)

# Set loop for each week of runs
for(j in 1:runs){
  # Build a table of results
  tab <- matrix(0, cust, 8)
  
  # Set initial time, wait time
  arr_time <- 0; beg_time <- 0; end_time <- 0; wait_time <- 0
  
  # Set up loop for every customer

# ---- PLACE CUSTOMER LOOP HERE ---- #
    
  # System results, column 7 is Wait, column 5 is service time, column 2 is arrival time
  total[j, ] <- matrix(t(c(cust,
                          (sum(tab[,7])/cust + sum(tab[,5])/cust)*(60/mean(tab[,2]))/60,
                          (sum(tab[,7])/cust)*(60/mean(tab[,2]))/60,
                          sum(tab[,7])/cust + sum(tab[,5])/cust,
                          sum(tab[,7])/cust)))
}
  
# Put header on table and name the header in the same line of code
colnames(total) <- t_name <- c("Customers", "L", "Lq", "W", "Wq")

```


Run the combined loops 
=============================================

```{r}
# Initialize results table
total <- matrix(0, runs, 5)

# Set loop for each week of runs
for(j in 1:runs){
  # Build a table of results
  tab <- matrix(0, cust, 7)
  
  # Set initial time, wait time
  arr_time <- 0; beg_time <- 0; end_time <- 0; wait_time <- 0
  
  # Set up loop for every customer
  for(i in 1:cust){
    # Minutes till next customer arrives to the system
    t_arr <- round(sample(e_arr, 1, replace = TRUE) * 60, 0)
  
    # Time at customer arrival
    arr_time <- arr_time + t_arr
    
    # Customer wait time
    wait_time <- ifelse(end_time > arr_time, end_time - arr_time, 0)
    
    # Set service begin time
    beg_time <- arr_time + wait_time 
    
    # Random service times for each customer
    t_ser <- round(sample(e_ser, 1, replace = TRUE) * 60, 0)
    
    # Set time service ends
    end_time <- beg_time + t_ser
  
    # Combine in a table
    tab[i, ] <- round(c(i, t_arr, arr_time, beg_time, t_ser, end_time, wait_time), 1) 
  }
  
  # System results, column 7 is Wait, column 5 is service time, column 2 is arrival time
  total[j, ] <- matrix(t(c(cust,
                          (sum(tab[,7])/cust + sum(tab[,5])/cust)*(60/mean(tab[,2]))/60,
                          (sum(tab[,7])/cust)*(60/mean(tab[,2]))/60,
                          sum(tab[,7])/cust + sum(tab[,5])/cust,
                          sum(tab[,7])/cust)))
}
  
# Put header on table and name the header in the same line of code
colnames(total) <- t_name <- c("Customers", "L", "Lq", "W", "Wq")

```


Table of key results
=============================================

<div style="float: left; width: 50%;">

```{r echo = F}
# queueing package mean results
(q_tab %>% 
  kable(caption = "Queueing package mean steady state results") %>%
  kable_styling(bootstrap_options = c("striped", "bordered")))

```

</div>

<div style="float: right; width: 50%;">

```{r echo = F}
# Make table
out <- round(rbind(apply(total, 2, mean), apply(total, 2, sd), 
             apply(total, 2, max), apply(total, 2, min)), 2)
rownames(out) <- c("mean", "sd", "max", "min")

# Print output
kable(out[,2:5], caption = "Custom queueing model simulation results") %>%
  kable_styling(bootstrap_options = c("striped", "bordered"))

```

</div>


Distribution of key variables
=============================================

+ **<span style="color: red;">red line</span>** is mean from simulation, **<span style="color: blue;">blue line</span>** is mean from `queueing` package

```{r echo = F, fig.width=12}
# Visualize results
p1 <- as.data.frame(total) %>% 
  ggplot(aes(x = total[,2])) + geom_histogram(bins = 20, fill = "white", color = "black") + 
  geom_vline(xintercept = mean(total[,2]), color = "red") +
  geom_vline(xintercept = as.numeric(q_tab[2]), color = "blue") + 
  labs(title = "Customers in the system per hour", x = "Customers") + theme_classic()

p2 <- as.data.frame(total) %>% 
  ggplot(aes(x = total[,3])) + geom_histogram(bins = 20, fill = "white", color = "black") + 
  geom_vline(xintercept = mean(total[,3]), color = "red") +
  geom_vline(xintercept = as.numeric(q_tab[3]), color = "blue") + 
  labs(title = "Customers in the queue per hour", x = "Customers") + theme_classic()

p3 <- as.data.frame(total) %>% 
  ggplot(aes(x = total[,4])) + geom_histogram(bins = 20, fill = "white", color = "black") + 
  geom_vline(xintercept = mean(total[,4]), color = "red") +
  geom_vline(xintercept = as.numeric(q_tab[4]), color = "blue") + 
  labs(title = "Customer time in the system", x = "Minutes") + theme_classic()

p4 <- as.data.frame(total) %>% 
  ggplot(aes(x = total[,5])) + geom_histogram(bins = 20, fill = "white", color = "black") + 
  geom_vline(xintercept = mean(total[,5]), color = "red") +
  geom_vline(xintercept = as.numeric(q_tab[5]), color = "blue") + 
  labs(title = "Customers wait time in the queue", x = "Minutes") + theme_classic()

(p1 + p2)/(p3 + p4)
```



Monte Carlo simulation
=============================================

+ The goal of simulation is to understand he performance of a system  

    - You are simulating a system because you cannot run experiments  
    
+ Monte Carlo simulation incorporates uncertainty  

    - Illuminates the impact of uncertainty on your system  
    
+ There are numerous ways to run Monte Carlo simulation  

    - But the general approach is the same  

+ We only several types of random variables  

    - Could have pulled observations from actual data  
    - treat like empirical distribution  
    

