---
title: "Week 3: Optimization under uncertainty"
subtitle: "Robust optimization and chance constraints"
author: "Gareth Green"
output: slidy_presentation
---

```{r echo = FALSE}
# Course: 5260 Math models for decision making
# Title: Week 3-Robust Optimization
# Purpose: Basic demonstration of Robust Optimization with lpSolveAPI
# Date: March 15, 2020
# Author: Gareth Green

```

```{r echo = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

```


Load packages
=============================================

```{r results = "hide"}
# Load packages
library(lpSolveAPI)
`%>%` <- magrittr::`%>%`
library(knitr)
library(kableExtra)

```


Homework #1, problem #4: Farming maximization  
=============================================

<div style="float: left; width: 50%;">

+ Maximize: 

    profit = 5G + 25V  

+ Subject to:  

    Materials 5G + 40V <= 1245  
    Water 4G +  8V <=  300  

```{r}
# Define parameters of the lp
h1p4 <- make.lp(0, 2)

```


```{r echo = FALSE, include = FALSE}
# I will run this in a hidden chunk to avoid the output
lp.control(h1p4, sense = "max")

```

</div>

<div style="float: right; width: 50%;">


```{r}
# Set up model
obj_fn <- c(5, 25)
set.objfn(h1p4, obj_fn)
add.constraint(h1p4, c(5, 40), "<=", 1245)
add.constraint(h1p4, c(4, 8), "<=", 300)

# Add row and column names
dimnames(h1p4) <- list(c("Input Mix", "Water"), 
                     c("Grains", "Veggies"))

# See the model we have set up
h1p4

```

```{r echo = FALSE, include = FALSE}
# Solve model
solve(h1p4)

# Make solution andsensitivity analysis table
# Primal values
ps <- get.primal.solution(h1p4) 

# Sensitivity analysis
obj_sa <- get.sensitivity.obj(h1p4)
rhs_sa <- get.sensitivity.rhs(h1p4)

n <- length(get.variables(h1p4))
m <- length(get.constr.type(h1p4))
ov <- paste0("Objective Value = ", ps[1])

sa_tab <- rbind(ps[2:(n + m + 1)], 
                c(round(rhs_sa$duals[1:m], 2), obj_fn),
                round(c(rhs_sa$dualsfrom[1:m],obj_sa$objfrom), 2),
                round(c(rhs_sa$dualstill[1:m],obj_sa$objtill), 2)) 
colnames(sa_tab) <- c(rownames(h1p4), colnames(h1p4))
rownames(sa_tab) <- c("Solution", "duals/coef", "Sens From", "Sens Till")  

# Substitute "inf" and "-inf" so tables are easier to read (coded by two classmates!)
sa_tab <- ifelse(sa_tab == -1.000e+30, "-inf", sa_tab)
sa_tab <- ifelse(sa_tab == 1.000e+30, "inf", sa_tab)

```


</div>

Recall the graphical representation
=============================================

```{r echo = FALSE}
# Plot feasible space
plot(h1p4)

```


Objective value and sensitivity analysis table
=============================================

<div style="float: left; width: 50%;">

```{r echo = FALSE}
# Print the table
kable(sa_tab, format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("striped", "bordered")) %>% 
  add_footnote(label = ov, notation = "none")

```


**What changes in the right hand side constraints (_b_'s) would change the solution**  

+ Duals indicate both the input mix and water constraints are binding  
  - Increasing water will increase profit by $0.62 per unit from 249 to 996 units
  - Increasing input mix will increase profit by $0.50 per unit from 375 to 1500 units
  - Increasing either resource will increase profits  
    - But not sure how crops (variables) will change  

</div>

<div style="float: right; width: 50%;">
            
**What changes in the objective function parameters (_c_'s) would change the solution**  

+ Will have same cropping pattern (variables) if profit per acre (c's) are **within** given ranges  
  - But profit will change  
  - For example, if increase profit of veggies from $25 to $39.99
    - Profit = $1244.97, G = 17 and V = 29

**What changes in the constraint parameters (_a_'s) would change the solution**  

+ This type of sensitivity analysis is not given by the output  
  - Ratio of $c$'s is 25/5 = 5  
  - Ratio of $a_i$ is 40/5 = 8, so $a_{Vi}$ decrease or $a_{Gi}$ increase so ratio < 5  
  - Ratio of $a_w$ is 8/4 = 2, so $a_{VW}$ increase or $a_{GW}$ decrease so ratio > 5   

**Let's try it to see**    

</div>


Robust Optimization-INDEPENDENT parameters  
=============================================

<div style="float: left; width: 55%;">

**Drought parameters** 

+ lower bound of c's: $c_G = 3$ and $c_V = 20$      
+ lower bound of water: $b_W = 200$  
+ upper bound of a's: $a_{GW} = 5$ and $a_{VW} = 10$  

+ maximize: 

    profit = 3G + 20V  

+ subject to:  

    Materials 5G + 40V <= 1245  
    Water 5G + 10V <= 200  

</div>

<div style="float: right; width: 45%;">

**Write down how you think the solution will change?**  

+ Objective value +/-  
+ Duals +/- 
+ Variables (acreage) +/-  

```{r}
# Define parameters of the lp
h1p4_ro_i <- make.lp(0, 2)

```

</div>

```{r echo = FALSE, include = FALSE}
# I will run this in a hidden chunk to avoid the output
lp.control(h1p4_ro_i, sense = "max")

```


```{r echo = TRUE, include = FALSE}
# Set up model
obj_fn_ro_i <- c(3, 20)
set.objfn(h1p4_ro_i, obj_fn_ro_i)
add.constraint(h1p4_ro_i, c(5, 40), "<=", 1245)
add.constraint(h1p4_ro_i, c(5, 10),"<=", 200)

# Add row and column names
dimnames(h1p4_ro_i) <- list(c("Input Mix", "Water"), 
                     c("Grains", "Veggies"))

# See the model we have set up
h1p4_ro_i

# Solve model
solve(h1p4_ro_i)

# Make solution andsensitivity analysis table
# Primal values
ps_ro_i <- get.primal.solution(h1p4_ro_i) 

# Sensitivity analysis
obj_sa_ro_i <- get.sensitivity.obj(h1p4_ro_i)
rhs_sa_ro_i <- get.sensitivity.rhs(h1p4_ro_i)

ov_ro_i <- paste0("Objective Value = ", ps_ro_i[1])

sa_tab_ro_i <- rbind(ps_ro_i[2:(n + m + 1)], 
             c(rhs_sa_ro_i$duals[1:m],obj_fn_ro_i),
             round(c(rhs_sa_ro_i$dualsfrom[1:m],obj_sa_ro_i$objfrom), 2),
             round(c(rhs_sa_ro_i$dualstill[1:m],obj_sa_ro_i$objtill), 2)) 
colnames(sa_tab_ro_i) <- c(rownames(h1p4_ro_i), colnames(h1p4_ro_i))
rownames(sa_tab_ro_i) <- c("Solution", "duals/coef", "Sens From", "Sens Till")

# Substitute "inf" and "-inf" so tables are easier to read (coded by two classmates!)
sa_tab_ro_i <- ifelse(sa_tab_ro_i == -1.000e+30, "-inf", sa_tab_ro_i)
sa_tab_ro_i <- ifelse(sa_tab_ro_i == 1.000e+30, "inf", sa_tab_ro_i)


```


Objective value and sensitivity analysis table
=============================================

<div style="float: left; width: 50%;">

```{r echo = FALSE}
# Print the table
kable(sa_tab_ro_i, format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("striped", "bordered")) %>% 
  add_footnote(label = ov_ro_i, notation = "none")

```

**How do results compare to your guess?**  

+ Objective function decrease  
+ Only water binding and the water dual increased  
+ Stop grains, reduce veggie

</div>

<div style="float: right; width: 50%;">

```{r echo = FALSE}
# Plot feasible space
plot(h1p4_ro_i)

```

</div>


Did robust optimization rules make sense?  
=============================================

Note: we applied the Robust Optimization rule WITHOUT considering the application  

+ **Does it make sense?**  

  - How would water use change during drought?  
    - Might be hotter so need more water, so yes that make sense  
  - How would c's and profit change?  
    - Grains price stay same due to world market, so not good assumption  
    - Veggie profit +/-? 
      - If mkt supply down, price increase, profit increase  
      - If water more due to heat could be more expensive, profit decrease
        - hard to say how profit per acre react
        - leave at usual price

+ So water use realistic, but profit parameters need to be studied


Robust Optimization-DEPENDENT parameters  
=============================================

<div style="float: left; width: 50%;">

**Drought parameters**  

+ Assume grains profit the same but veggies profit increase due to reduced supply  
  - Putting in fixed values derived from functional relationship

+ fixed functional value of c's: $c_G = 5$ and $c_V = 25$      
+ lower bound of water: $b_W = 200$  
+ upper bound of a's: $a_{GW} = 5$ and $a_{VW} = 10$  

+ refine c's: $c_G = 5$ and $c_V = 25$    
  - Base value of c for grain since world market  
  - Higher profit for veggie since lower production  
    - However, this is not conservative approach  
+ lower bound of water: $b_W = 200$  
+ upper bound of a's: $a_{GW} = 5$ and $a_{VW} = 10$     
  - since use more water if hotter in drought  

**Why is this robust optimization with dependent parameters?**  

+ Have changed parameters based on case rather than just follow rules  
+ Have simplified to use linear programming  
  - Not have a's or c's as function of water or temperature  
    
</div>

<div style="float: right; width: 50%;">
    
+ maximize: 

    profit = 5G + 25V  

+ subject to:  

    Materials 5G + 40V <= 1245  
    Water 5G +  10V <=  200  

**Write down how you think the solution will change?**  

+ Objective value +/-  
+ Duals +/- 
+ Variables (acreage) +/-  

```{r}
# Define parameters of the lp
h1p4_ro_d <- make.lp(0, 2)

```

</div>

```{r echo = FALSE, include = FALSE}
# I will run this in a hidden chunk to avoid the output
lp.control(h1p4_ro_d, sense = "max")

```


```{r echo = TRUE, include = FALSE}
# Set up model
obj_fn_ro_d <- c(5, 25)
set.objfn(h1p4_ro_d, obj_fn_ro_d)
add.constraint(h1p4_ro_d, c(5, 40), "<=", 1245)
add.constraint(h1p4_ro_d, c(5, 10),"<=", 200)

# Add row and column names
dimnames(h1p4_ro_d) <- list(c("Input Mix", "Water"), 
                        c("Grains", "Veggies"))

# See the model we have set up
h1p4_ro_d

# Solve model
solve(h1p4_ro_d)

# Make solution andsensitivity analysis table
# Primal values
ps_ro_d <- get.primal.solution(h1p4_ro_d) 

# Sensitivity analysis
obj_sa_ro_d <- get.sensitivity.obj(h1p4_ro_d)
rhs_sa_ro_d <- get.sensitivity.rhs(h1p4_ro_d)

ov_ro_d <- paste0("Objective Value = ", ps_ro_d[1])

sa_tab_ro_d <- rbind(ps_ro_d[2:(n + m + 1)], 
             c(rhs_sa_ro_d$duals[1:m],obj_fn_ro_d),
             round(c(rhs_sa_ro_d$dualsfrom[1:m],obj_sa_ro_i$objfrom), 2),
             round(c(rhs_sa_ro_d$dualstill[1:m],obj_sa_ro_i$objtill), 2)) 
colnames(sa_tab_ro_d) <- c(rownames(h1p4_ro_d), colnames(h1p4_ro_d))
rownames(sa_tab_ro_d) <- c("Solution", "duals/coef", "Sens From", "Sens Till")      

# Substitute "inf" and "-inf" so tables are easier to read (coded by two classmates!)
sa_tab_ro_d <- ifelse(sa_tab_ro_d == -1.000e+30, "-inf", sa_tab_ro_d)
sa_tab_ro_d <- ifelse(sa_tab_ro_d == 1.000e+30, "inf", sa_tab_ro_d)

```

Objective value and sensitivity analysis table
=============================================

<div style="float: left; width: 50%;">

```{r echo = FALSE}
# Print the table
kable(sa_tab_ro_d, format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("striped", "bordered")) %>% 
  add_footnote(label = ov_ro_d, notation = "none")

```


**How do results compare to your guess?**  

+ Objective function decrease, but not by as much as before (was $400)  
+ Only water binding and the water dual increased  
+ Stop grains, reduce veggie

</div>

<div style="float: right; width: 50%;">

```{r echo = FALSE}
# Plot feasible space
plot(h1p4_ro_d)

```

</div>


Chance Constraint 
=============================================

<div style="float: left; width: 50%;">

**Drought parameters**  

+ use refined c's: $c_{GW} = 5$ and $c_{VW} = 25$    
  - Base profit of c for grain and for veggie (since not know how change)
+ upper bound of a's: $a_{GW} = 5$, $a_{VW} = 10$     
  - since use more water  

Suppose distribution of water is $\mu = 300$ and $\sigma = 50$  

  + Lower bound of water: $b_{water}\le\mu-K_{\alpha}*\sigma=300-K_{\alpha}*50$  
  + Use 90% confidence so $\alpha = 0.1$
  + But is a two-tail so only get 0.05 in lower tail

</div>

<div style="float: right; width: 50%;">

+ maximize: 

    profit = 5G + 25V  

+ subject to:  

    Materials 5G + 40V <= 1245  
    Water 5G +  10V <=  300 - $K_{\alpha}$ * 50  

**Write down how you think the solution will change?**  

+ Objective value +/-  
+ Duals +/- 
+ Variables (acreage) +/-  

```{r}
# Define parameters of the lp
h1p4_cc <- make.lp(0, 2)

```

</div>

```{r echo = FALSE, include = FALSE}
# I will run this in a hidden chunk to avoid the output
lp.control(h1p4_cc, sense = "max")

```


```{r echo = TRUE, include = FALSE}
# Set up model with chance constraints on veggie profit and water use
obj_fn_cc <- c(5, 25) # qnorm gives value from normal dist
set.objfn(h1p4_cc, obj_fn_cc)
add.constraint(h1p4_cc, c(5, 40), "<=", 1245) # input mix unchanged in drought
add.constraint(h1p4_cc, c(5, 10), "<=", qnorm(0.05, mean = 300, sd = 50))

# Add row and column names
dimnames(h1p4_cc) <- list(c("Input Mix", "Water"), 
                     c("Grains", "Veggies"))

# See the model we have set up
h1p4_cc

# Solve model
solve(h1p4_cc)

# Make solution andsensitivity analysis table
# Primal values
ps_cc <- get.primal.solution(h1p4_cc) 

# Sensitivity analysis
obj_sa_cc <- get.sensitivity.obj(h1p4_cc)
rhs_sa_cc <- get.sensitivity.rhs(h1p4_cc)

ov_cc <- paste0("Objective Value = ", ps_cc[1])

sa_tab_cc <- rbind(round(ps_cc[2:(n + m + 1)], 2), 
             round(c(rhs_sa_cc$duals[1:m],obj_fn_cc), 2),
             round(c(rhs_sa_cc$dualsfrom[1:m],obj_sa_cc$objfrom), 2),
             round(c(rhs_sa_cc$dualstill[1:m],obj_sa_cc$objtill), 2)) 
colnames(sa_tab_cc) <- c(rownames(h1p4_cc), colnames(h1p4_cc))
rownames(sa_tab_cc) <- c("Solution", "duals/coef", "Sens From", "Sens Till")      

# Substitute "inf" and "-inf" so tables are easier to read (coded by two classmates!)
sa_tab_cc <- ifelse(sa_tab_cc == -1.000e+30, "-inf", sa_tab_cc)
sa_tab_cc <- ifelse(sa_tab_cc == 1.000e+30, "inf", sa_tab_cc)

```

Objective value and sensitivity analysis table
=============================================

<div style="float: left; width: 50%;">

```{r echo = FALSE}
# Print the table
kable(sa_tab_cc, format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("striped", "bordered")) %>% 
  add_footnote(label = ov_cc, notation = "none")

```


**How do results compare to your guess?**  

+ Objective function decrease  
  - But not by as much since since lower bound 218 > 200  
+ Only water binding and the water dual increased  
+ Stop grains, reduce veggie

</div>

<div style="float: right; width: 50%;">

```{r echo = FALSE}
# Plot feasible space
plot(h1p4_cc)

```

</div>


