---
title: "Week 3"
subtitle: "Transportation or Assignment problem with dummy destination reading in data from csv and looping"
author: "Gareth Green"
output: slidy_presentation
---

```{r echo = FALSE}
# Title: Week 3-Transportation or Assignment Problem with dummy destination
# Purpose: Demonstrate lpSolveAPI on pg353 Prototype reading in data from csv
# Date: March 15, 2020
# Author: Gareth Green

```

```{r echo = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

```

```{r echo = FALSE}
# Load packages
library(lpSolveAPI)
`%>%` <- magrittr::`%>%`
library(knitr)
library(kableExtra)

```


Backgound: Transportation or Assignment problem, pg 353
=============================================

<div style="float: left; width: 50%;">

+ Transporation and Assignment problems need to maintain balance of inputs and outputs  

  - For transportation problems need to balance quantity (supply and demand)
  - For assignment problem need to balance tasks and assignees
  - If the probelm is not balanced need to set up dummy to maintain balance
    - excess supply (source or assignee) need dummy demand (destination or task), like in this problem
    - excess demand (destination or task) need dummy supply (source or assignee)

+ We will solve the probelm two ways, as Transportation and Assignment

</div>
<div style="float: right; width: 50%;">

+ Assigning production of products to different plants based on cost and capacity

  - Read through problem for details
  
+ Option 1: Treat like Transportation problem

  - **production splitting**, a plant can produce more than one product

+ Option 2: Treat like Assigment problem

  - **No production splitting**, a plant can only produce all  of one or two products

+ Again, read 353 - 6 for details

</div>


Option 1: Set up the model from csv file and loops
=============================================

```{r}
# Read in data file of costs and rhs
mat <- read.csv("W3-4-2_as_9_3_ex2_opt1.csv", header = FALSE)

# Seperate out objective function
m <- nrow(mat) - 1
n <- ncol(mat) - 1
obj_fn <- unlist(as.data.frame(t(mat[1:m, 1:n])))

# Make model
as_9_3_ex2_opt1 <- make.lp(0, n * m)

# Set up model
set.objfn(as_9_3_ex2_opt1, obj_fn)

# Set up row constraints
for(i in 0:(m - 1)){ 
  add.constraint(as_9_3_ex2_opt1, c(rep(0, i * n), rep(1, n), rep(0, (m-1-i) * n)), "=", mat[(i+1),(n+1)])
  i <- i + 1
}

for(j in 0:(n-1) ){ 
  add.constraint(as_9_3_ex2_opt1, (rep(c(rep(0,j),1,rep(0, n-1-j)),m)), "=", mat[(m+1),(j+1)])
  j <- j + 1
}

# Does it have the right dimmensions?
as_9_3_ex2_opt1

# Write to view the algebraic formulation
write.lp(as_9_3_ex2_opt1, "W3-4-2_as_9_3_ex2_opt1.lp",type = 'lp')

```


Option 1: Solve and get results Transportation set up
=============================================

```{r}
# Solve the model, if this return 0 an optimal solution is found
solve(as_9_3_ex2_opt1)

```

```{r echo = FALSE, include = FALSE}
### Sensitivity Analysis
ps <- get.primal.solution(as_9_3_ex2_opt1)
obj_sa <- get.sensitivity.obj(as_9_3_ex2_opt1)
rhs_sa <- get.sensitivity.rhs(as_9_3_ex2_opt1)

nv <- length(get.variables(as_9_3_ex2_opt1))
mc <- length(get.constr.type(as_9_3_ex2_opt1))
ov <- paste0("Objective Value = ", ps[1])

sa_tab <- rbind(ps[2:(nv + mc + 1)], 
                round(c(rhs_sa$duals[1:mc], obj_fn), 2),
                round(c(rhs_sa$dualsfrom[1:mc],obj_sa$objfrom), 2),
                round(c(rhs_sa$dualstill[1:mc],obj_sa$objtill), 2)) 
colnames(sa_tab) <- c(rownames(as_9_3_ex2_opt1), colnames(as_9_3_ex2_opt1))
rownames(sa_tab) <- c("solution", "duals/coef", "Sens From", "Sens Till")      

# Substitute "inf" and "-inf" so tables are easier to read (coded by two classmates!)
sa_tab <- ifelse(sa_tab == -1.000e+30, "-inf", sa_tab)
sa_tab <- ifelse(sa_tab == 1.000e+30, "inf", sa_tab)

```

```{r echo = FALSE}
# Print the table
kable(sa_tab, format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("striped", "bordered")) %>% 
  add_footnote(label = ov, notation = "none")

```


Option 2: Set up the model from csv file and loops
=============================================

```{r}
# Read in data file of costs and rhs
mat <- read.csv("W3-4-2_as_9_3_ex2_opt2.csv", header = FALSE)

# Seperate out objective function
m <- nrow(mat) - 1
n <- ncol(mat) - 1
obj_fn <- unlist(as.data.frame(t(mat[1:m, 1:n])))

# Make model
as_9_3_ex2_opt2 <- make.lp(0, n * m)

# Set up model
set.objfn(as_9_3_ex2_opt2, obj_fn)

# Set up row constraints
for(i in 0:(m - 1)){ 
  add.constraint(as_9_3_ex2_opt2, c(rep(0, i * n), rep(1, n), rep(0, (m-1-i) * n)), "=", mat[(i+1),(n+1)])
  i <- i + 1
}

for(j in 0:(n-1) ){ 
  add.constraint(as_9_3_ex2_opt2, (rep(c(rep(0,j),1,rep(0, n-1-j)),m)), "=", mat[(m+1),(j+1)])
  j <- j + 1
}

# Does it have the right dimmensions?
as_9_3_ex2_opt2

# Write to view the algebraic formulation
write.lp(as_9_3_ex2_opt2, "W3-4-2_as_9_3_ex2_opt2.lp",type = 'lp')

```


Solve and get results of Assignment set up, Option 2
=============================================

```{r}
# Solve the model, if this return 0 an optimal solution is found
solve(as_9_3_ex2_opt2)

```

```{r echo = FALSE, include = FALSE}
### Sensitivity Analysis
ps <- get.primal.solution(as_9_3_ex2_opt2)
obj_sa <- get.sensitivity.obj(as_9_3_ex2_opt2)
rhs_sa <- get.sensitivity.rhs(as_9_3_ex2_opt2)

nv <- length(get.variables(as_9_3_ex2_opt2))
mc <- length(get.constr.type(as_9_3_ex2_opt2))
ov <- paste0("Objective Value = ", ps[1])

sa_tab <- rbind(ps[2:(nv + mc + 1)], 
                round(c(rhs_sa$duals[1:mc], obj_fn), 2),
                round(c(rhs_sa$dualsfrom[1:mc],obj_sa$objfrom), 2),
                round(c(rhs_sa$dualstill[1:mc],obj_sa$objtill), 2)) 
colnames(sa_tab) <- c(rownames(as_9_3_ex2_opt2), colnames(as_9_3_ex2_opt2))
rownames(sa_tab) <- c("solution", "duals/coef", "Sens From", "Sens Till")      

# Substitute "inf" and "-inf" so tables are easier to read (coded by two classmates!)
sa_tab <- ifelse(sa_tab == -1.000e+30, "-inf", sa_tab)
sa_tab <- ifelse(sa_tab == 1.000e+30, "inf", sa_tab)

```

```{r echo = FALSE}
# Print the table
kable(sa_tab, format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("striped", "bordered")) %>% 
  add_footnote(label = ov, notation = "none")

```


