---
title: "Technical Appendix - Exhibition"
author: "Renato Albolea"
date: "6/3/2020"
output: 
  html_document: 
    keep_md: yes
    toc: yes
    toc_float: true
    toc_depth: 4
    code_download: yes
---
```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
# if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

library(tidyverse)
library(magrittr)

library(knitr)
library(kableExtra)
library(formattable)
library(patchwork)
library(scales)
library(lpSolveAPI)
library(here)
```


```{r echo = FALSE}
# Function to generate the sensitivity table
sensitivity_table <- function(model){
  # Make solution/sensitivity analysis table
  # Get primal solution
  ps <- get.primal.solution(model)
  
  
  # Get sensitivity analysis
  obj_sa <- get.sensitivity.obj(model)
  rhs_sa <- get.sensitivity.rhs(model)
  
  n <- length(get.variables(model))
  m <- length(get.constr.type(model))
  ov <- paste0("Objective Value = ", comma(ps[1], accuracy = 0.01))
  
  sa_tab <- rbind(round(ps[2:(n + m + 1)], 2), 
                  c(round(rhs_sa$duals[1:m], 2), obj_fn),
                  round(c(rhs_sa$dualsfrom[1:m],obj_sa$objfrom), 2),
                  round(c(rhs_sa$dualstill[1:m],obj_sa$objtill), 2)) 
  colnames(sa_tab) <- c(rownames(model), colnames(model))
  rownames(sa_tab) <- c("solution", "duals/coef", "Sens From", "Sens Till")      
  
  # Substitute "inf" and "-inf" so tables are easier to read (coded by two classmates!)
  sa_tab <- ifelse(sa_tab <= -1.000e+30, "-inf", sa_tab)
  sa_tab <- ifelse(sa_tab >= 1.000e+30, "inf", sa_tab)
  
  # Print the table
  kable(sa_tab, format.args = list(big.mark = ",")) %>%
    kable_styling(bootstrap_options = c("striped", "bordered")) %>% 
    add_footnote(label = ov, notation = "none")
}
  
```

```{r echo = FALSE}
# Function to generate the sensitivity table
result_table <- function(model){
  # Make solution/sensitivity analysis table
  # Get primal solution
  ps <- get.primal.solution(model)
  
  n <- length(get.variables(model))
  m <- length(get.constr.type(model))
  ov <- paste0("Objective Value = ", comma(ps[1], accuracy = 0.01))

  sa_tab <- rbind(round(ps[2:(n + m + 1)], 2))
  colnames(sa_tab) <- c(rownames(model), colnames(model))
  rownames(sa_tab) <- c("solution")
  
  # Substitute "inf" and "-inf" so tables are easier to read (coded by two classmates!)
  sa_tab <- ifelse(sa_tab <= -1.000e+30, "-inf", sa_tab)
  sa_tab <- ifelse(sa_tab >= 1.000e+30, "inf", sa_tab)
  
  # Print the table
  kable(sa_tab, format.args = list(big.mark = ",")) %>%
    kable_styling(bootstrap_options = c("striped", "bordered")) %>% 
    add_footnote(label = ov, notation = "none")
}
  
```

# Objective:  

The Guggenheim Museum is interested in curating a new exhibit. This technical appendix brings support for 3 main proposal of which pieces should be displayed and the total cost of each proposal.  
\  

# General Restrictions:  

+ Must include:
  1. Only one collage  
  1.  
     a. At least one wire mesh sculpture displayed if a computer-generated drawing is displayed __and__  
     b. At least one computer-generated drawing displayed if a wire mesh sculpture is
displayed  
  1. At least one photo-realistic painting 
  1. At least one cubist painting  
  1. At least one expressionist painting  
  1. At least one watercolor painting  
  1. At least one oil painting  
  1. The number of paintings to be no greater than three times the number of other art forms  
  1. All Valentino's paintings   
  1. Both of Hellen's paintings  
  1. Display as many pieces from David as from Enrique  
  1. At least one piece from each of them (David and Enrique)  
  1. Only display one piece from Rick, if any at all.  
  1. At least one piece from a female artist for every three pieces included from a male artist.  
  1. Either one or both of the pieces “Aging Earth” and “Wasted Resources.”  
  1. At least one piece by Stuart  
  1. At least one of the following pieces: “Beyond,” “Who Has Control,” “Where are we now?” and  “Pioneers.”   
\  
\  
+ The museum only has enough space for:  
  18. Four sculptures  
  1. Around 20 paintings, collages, and drawings.  
\  
\  
  20. If “Narcissism” is displayed, “Reflection” should also be displayed  
\  
\  

```{r}
general_restrictions <- function(model){
  # Add constraints in the order presented in the 'General Restrictions'
  add.constraint(model, c(art_col$Form=="Collage"), "=", 1)
  add.constraint(model, -c((art_col$Form=="Wire")*(1/sum(art_col$Form=="Wire")))+c(art_col$Form=="Computer"), ">=", 0)
  add.constraint(model, -c((art_col$Form=="Computer")*(1/sum(art_col$Form=="Computer")))+c(art_col$Form=="Wire"), ">=", 0)
  add.constraint(model, c(art_col$Form=="Photo"), ">=", 1)
  add.constraint(model, c(art_col$Form=="Cubist"), ">=", 1)
  add.constraint(model, c(art_col$Form=="Expressionist"), "=", 1)
  add.constraint(model, c(art_col$Form=="Watercolor"), ">=", 1)
  add.constraint(model, c(art_col$Form=="Oil"), ">=", 1)
  add.constraint(model, c(art_col$Type=="Painting")-3*c(art_col$Type!="Painting"), "<=", 0)
  add.constraint(model, c(art_col$Artist=="Valentino"), "=", sum(art_col$Artist=="Valentino"))
  add.constraint(model, c(art_col$Artist=="Helen"), "=", sum(art_col$Artist=="Helen"))
  add.constraint(model, c(art_col$Artist=="David")-c(art_col$Artist=="Enrique"), "=", 0)
  add.constraint(model, c(art_col$Artist=="David"), ">=", 1) # I don't need to use Enrique
  add.constraint(model, c(art_col$Artist=="Rick"), "<=", 1)
  add.constraint(model, c(art_col$Gender=="M")-3*c(art_col$Gender=="F"), "<=", 0)
  add.constraint(model, c((art_col$Piece=="Aging Earth" |
                            art_col$Piece=="Wasted Resources")), ">=", 1)
  add.constraint(model, c(art_col$Artist=="Stuart"), ">=", 1)
  add.constraint(model, c((art_col$Piece=="Beyond" |
                              art_col$Piece=="Who Has Control?" |
                              art_col$Piece=="Where are we now?" |
                              art_col$Piece=="Pioneers" )), ">=", 1)
  add.constraint(model, c(art_col$Type=="Sculpture"), "<=", 4)
  add.constraint(model, c(art_col$Type!="Sculpture"), "<=", 20)
  add.constraint(model, c(art_col$Piece=="Narcissism") - c(art_col$Piece=="Reflection"), "<=", 0)
  
  constraints_name <- c("1_coll",
                      "at_l_1_w_if_c_d",
                      "at_l_1_d_if_c_w",
                      "at_l_1_pht",
                      "at_l_1_cub",
                      "at_l_1_expt",
                      "at_l_1_wcol",
                      "at_l_1_oil",
                      "paint<=3non",
                      "valentino",
                      "helen",
                      "David=Enrique",
                      "at_l_1_Darcy",
                      "Rick<=1",
                      "3Fem=Male",
                      "AErth_a_o_WR",
                      "at_l_1_Stuart",
                      "at_l_1_4_pieces",
                      "max_4_scul",
                      "max_20_n_scul",
                      "Narc_with_Refl")
  return(constraints_name)
}
```


# Proposal 1: Limited Budget  

+ Objective: Maximize the number of pieces displayed in the exhibit without exceeding the budget($4 million)  

```{r import_data, message=FALSE}
# Import Art Collection
art_col <- read_csv(here("Final_Project", "Data", "Data.csv"))

# Include Sex

art_col %<>% 
  rename(Desc = `Description of Piece`) %>% 
  mutate(Form = case_when(str_detect(Desc, 'wire mesh') ~ 'Wire',
                          str_detect(Desc, 'computer') ~ 'Computer',
                          str_detect(Desc, 'pen-and-ink') ~ 'Ink',
                          str_detect(Desc, 'sculpture') ~ 'Sculpture',
                          str_detect(Desc, 'collage') ~ 'Collage',
                          str_detect(Desc, 'watercolor') ~ 'Watercolor',
                          str_detect(Desc, 'oil') ~ 'Oil',
                          str_detect(Desc, 'photo') ~ 'Photo',
                          str_detect(Desc, 'cubist') ~ 'Cubist',
                          str_detect(Desc, 'expressionist') ~ 'Expressionist'),
         Type  = case_when(str_detect(Desc, 'sculpture') ~ 'Sculpture',
                           str_detect(Desc, 'painting') ~ 'Painting',
                           str_detect(Desc, 'collage') ~ 'Collage',
                           str_detect(Desc, 'drawing') ~ 'Drawing'),
         Name = str_c(Artist,
                      str_sub(Piece, start = 1L, end = 3L), 
                      str_sub(Form, start = 1L, end = 3L),
                      Gender,
                      str_sub(Type, start = 1L, end = 3L),
                      sep="_")
         )
```

```{r model_prop_1}
n_var <- nrow(art_col)

# Set up the number of pieces maximization
prop_1 <- make.lp(0, n_var)

```


```{r echo = TRUE, results=FALSE}
# Run as maximum
lp.control(prop_1, sense = "max")

```

```{r model_prop_1_cont}
# Build objective function
obj_fn <- c(rep(1,n_var))

set.objfn(prop_1, obj_fn)

# Force binary variables on the solution 
set.type(prop_1, 1:n_var, "binary")

# add general Restrictions
constraints_name <- general_restrictions(prop_1)

# 
add.constraint(prop_1, c(art_col$Price), "<=", 4000000)

# # Add row and column names
constraints_name <- c(constraints_name,
                      "Max_Budget")

var_interest_name <- c(art_col$Name)

dimnames(prop_1) <- list(constraints_name, var_interest_name)

# Write to see named algebraic formulation
write.lp(prop_1, "Proposal_1.lp",type = 'lp')

# Solve the model, if this return 0 an optimal solution is found
solve(prop_1)
```




```{r}
result_table(prop_1)
```


With the proposal above we would be able to display  `r get.objective(prop_1)` art pieces. The selected pieces are:

```{r}
art_col %>% 
  filter(get.variables(prop_1)==TRUE) %>%  
  select(!(Form:Name)) %>% 
  kable(format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("striped", "bordered")) %>% 
  cat(., file = "../Final_Project/Tables/prop_1.html")

htmltools::includeHTML("../Final_Project/Tables/prop_1.html")
```



# Proposal 2: Minimum of 20 Pieces  

+ Objective: Minimize the cost of exibit at least 20 pieces with the same General Restrictions.   


```{r model_prop_2}
n_var <- nrow(art_col) # we will use One addtional variable for constraint 2 a/b

# Set up the number of pieces
prop_2 <- make.lp(0, n_var)

```

```{r echo = TRUE, results=FALSE}
# Run as minimum
lp.control(prop_2, sense = "min")

```


```{r}
# Build objective function
obj_fn <- c(art_col$Price)

set.objfn(prop_2, obj_fn)

# Force binary variables on the solution 
set.type(prop_2, 1:n_var, "binary")

# add general Restrictions
constraints_name <- general_restrictions(prop_2)

# 
add.constraint(prop_2, rep(1,n_var), ">=", 20)

# # Add row and column names
constraints_name <- c(constraints_name,
                      "At_l_20_Pieces")

var_interest_name <- c(art_col$Name)

dimnames(prop_2) <- list(constraints_name, var_interest_name)

# Write to see named algebraic formulation
write.lp(prop_2, "Proposal_2.lp",type = 'lp')

# Solve the model, if this return 0 an optimal solution is found
solve(prop_2)
```

```{r}
result_table(prop_2)
```


With the proposal above we would be able to display  `r sum(get.variables(prop_2))` art pieces at a cost of $`r comma(get.objective(prop_2)/1000000, accuracy=0.01)` million. The selected pieces are:

```{r}
art_col %>% 
  filter(get.variables(prop_2)==TRUE) %>%  
  select(!(Form:Name)) %>% 
  kable(format.args = list(big.mark = ",")) %>%
    kable_styling(bootstrap_options = c("striped", "bordered"))%>% 
  cat(., file = "../Final_Project/Tables/prop_2.html")

htmltools::includeHTML("../Final_Project/Tables/prop_2.html")
```



# Proposal 3: Adjusting Pieces Cost  

+ Objective: Considering the possible price's adjustments, what is the impact on the cost of Proposal 2?  

```{r}
# number of simulations
n <- seq(1:1000)

# Generate random variable with uniform distribution
set.seed(1234)

# Make the function
simulation <- function(n_simulation){
  price_factor <- runif(n_var, 0.8, 1.2)
  
  # Define parameters of the lp
  prop_3 <- make.lp(0, n_var)
  set.type(prop_3, 1:n_var, "binary")
  set.objfn(prop_3, c(price_factor*art_col$Price))
  
  # add general Restrictions
  general_restrictions(prop_3)
  # 
  add.constraint(prop_3, rep(1,n_var), ">=", 20)

  # Solve model
  solve(prop_3)

  # Output
  return(get.objective(prop_3))
}

res_simu_prop_3 <- (sapply(n, simulation)) # t() to traspose for correct dimmensions

mean_simu <- mean(res_simu_prop_3)/1000000
sd_simu <- sd(res_simu_prop_3)/1000000

# Make table
out_prop_3 <- round(
  rbind(
    c(mean_simu, (mean_simu*1000000/get.objective(prop_2)-1)*100),
    c(sd_simu, 0), 
    c(mean_simu-qnorm(0.975)*sd_simu, ((mean_simu-qnorm(0.975)*sd_simu)*1000000/get.objective(prop_2)-1)*100),
    c(mean_simu+qnorm(0.975)*sd_simu, ((mean_simu+qnorm(0.975)*sd_simu)*1000000/get.objective(prop_2)-1)*100),
    c(max(res_simu_prop_3)/1000000, (max(res_simu_prop_3)/get.objective(prop_2)-1)*100),
    c(min(res_simu_prop_3)/1000000, (min(res_simu_prop_3)/get.objective(prop_2)-1)*100)
    ), 2)
rownames(out_prop_3) <- c("mean", "sd", "Conf Interval 2.5%","Conf Interval 97.5%","max", "min")
colnames(out_prop_3) <- c("Simulated Budget ", "Impact on the original budget(%)")

# Print output
kable(out_prop_3, caption = "Loop results") %>%
  kable_styling(bootstrap_options = c("striped", "bordered"))


as.data.frame(res_simu_prop_3) %>% 
  ggplot(aes(x=res_simu_prop_3/1000000)) +
  geom_histogram(bins = 25, fill = "white", color = "black")+
  geom_vline(xintercept = mean_simu, color = "red")+
  geom_vline(xintercept = get.objective(prop_2)/1000000, color = "blue")+
  geom_vline(xintercept = mean_simu-qnorm(0.975)*sd_simu, color = "darkgreen")+
  geom_vline(xintercept = mean_simu+qnorm(0.975)*sd_simu, color = "darkgreen")+
  labs(title = "", x = "Budget Range") + theme_classic()
```



## Save Budget Values
```{r}
results <- as.data.frame(cbind(get.objective(prop_2)/1000000,
                               mean_simu-qnorm(0.975)*sd_simu,
                               mean_simu+qnorm(0.975)*sd_simu,
                               mean_simu
                               ))
colnames(results) <- c("prop_2", 
                       "prop_3_low_ci",
                       "prop_3_high_ci",
                       "prop_3_mean")
write.csv(results, file = "../Final_Project/Data/results.csv", row.names = FALSE)
```



  